%
%	The ARexx User's Manual
%
% Copyright © 1987 by William S. Hawes.  All Rights Reserved.
%
\ifx\radeye\fmtversion\subdoc\else\input pmac \fi
%
%   The Title Page
%
\null
\centerline{\bigrm \DOC}
\vskip 10pt
\centerline{\cfont Version 1.0}
\vskip 144pt
\centerline{\bigit The REXX Language for the Amiga}
\vskip 144pt
\centerline{\bf Copyright \copyright~~1987 William S. Hawes}
\centerline{\bf All Rights Reserved}
\page
%
%   Copyright Notices
%
\leftline{\bf Copyright Notice}
\noindent
\AR\ software and documentation are Copyright \copyright 1987 by 
\hbox{William S. Hawes}.
No part of the software or documentation may be reproduced,
transmitted, translated into other languages, posted to a network,
or distributed in any way without the express written permission of the author.
\vb
\leftline{\bf Disclaimer}
\noindent
This product is offered for sale ``as is'' with no representation of fitness
for any particular purpose.
The user assumes all risks and responsibilities related to its use.
The material within is believed to be accurate, 
but the author reserves the right to make changes to the software or
documentation without notice.
\vb
\leftline{\bf {Distribution}}
\noindent
\AR\ software and documentation are available from:
\vb
\centerline{\vbox{\hbox{William S. Hawes}
                  \hbox{P.O. Box 308}
                  \hbox{Maynard, MA~~01754}
                  \hbox{(508) 568-8695}}}
\noindent
Please direct orders or inquiries about this product to the above address.
Site licenses are available; write for further information.
\vb
\leftline{\bf About ...}
\noindent
\AR\ was developed on an Amiga 1000 computer with 512K bytes of memory
and two floppy disk drives.
The language prototype was developed in C using Lattice C,
and the production version was written in assembly-language using the
Metacomco Assembler.
The documention was created using the TxEd editor,
and was set in \TeX\ using Amiga\TeX.
This is a 100\% Amiga product.
\vskip 144pt
\leftline{\bf Trademarks}
\noindent
Amiga, Amiga WorkBench, and Intuition are trademarks of Commodore-Amiga, Inc.
\vfil

\thatsit
